---
title: |
  | **DS 223: Marketing Analytics**  
  | Homework 3 - Survival Analysis
author: "Alexander Shahramanyan"
date: "April 28, 2024"
output: 
  pdf_document:
    latex_engine: xelatex
classoption: a4paper
urlcolor: blue
header-includes:
  - \usepackage{titling}
  - \setlength{\droptitle}{-2cm}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, echo = TRUE, dev='pdf')
```


```{r echo=F}
# Load the libraries
library(survival)
library(texreg)
library(ggplot2)
library(patchwork)
library(tibble) 
```

To build AFT models, we first need to load the data. We will be working with the Telco Customer Churn dataset, which has the following columns:
- ID: subscriber's ID
- region: region code
- tenure: lifetime (in months)
- age: subscriber’s age
- marital: subscriber’s marital status
- address: number of years living in the same address
- income: subscriber’s annual income (K)
- ed: subscriber's education level
- retire: retired (Yes/No)
- gender: subscriber's gender (Male/Female)
- voice: voice service (Yes/No)
- internet: internet service (Yes/No)
- forward: call forwarding (Yes/No)
- custcat: customer category
- churn: whether the customer churned (Yes/No)

```{r}
# Read the CSV file
telco <- read.csv("telco.csv")
telco$churn = ifelse(telco$churn=='Yes', 1, 0)

head(telco)
```
Now, we will build basic models (intercept-only) with all the different distributions available in `survreg` package.

```{r}
surv_obj = Surv(time=telco$tenure, event=telco$churn)
reg_models <- list()

for(distribution in names(survreg.distributions)){
    # get the regression model
    reg_m = survreg(formula=surv_obj~1, dist=distribution)
    
    # print the summary
    # summary(reg_m)
    
    # add reg_m to reg_models
    reg_models[[distribution]] <- reg_m
}
```

As we have the models now, let's visualize the probability of churn during customer lifetime using the models in different plots and have an initial look at them.


```{r, fig.height=12, fig.width=10}
# Initialize an empty list for storing plots
plot_list <- list()

for (distribution in names(survreg.distributions)) {
    reg_m <- reg_models[[distribution]]
    
    probs <- seq(.1, .9, length=9)
    pred <- predict(reg_m, type="quantile", p=1-probs, newdata=data.frame(1))
    
    df <- data.frame(Time=pred, Probabilities=probs)

    # Generate the plot for current distribution
    p <- ggplot(df, aes(x = Time, y = Probabilities)) +
        geom_line() +
        labs(title = paste("Survival Curve |", distribution),
             x = "Survival Time",
             y = "Survival Probability")
    
    # Store the plot in the list
    plot_list[[distribution]] <- p
}

# Combine the plots into a grid (4x3) and leave the last two positions blank
plot_grid <- wrap_plots(plot_list, nrow = 4, ncol = 3) +
             plot_spacer() + plot_spacer()

# Print the combined plot grid
print(plot_grid)
```

As we can see, there are indeed some differences between the models. We can plot all the model curves in one graph to be able to compare the models.

```{r, fig.height=8, fig.width=8}
# Initialize an empty data frame for storing combined data
combined_df <- data.frame()

for (distribution in names(survreg.distributions)) {
    reg_m <- reg_models[[distribution]]
    
    probs <- seq(.1, .9, length=9)
    pred <- predict(reg_m, type="quantile", p=1-probs, newdata=data.frame(1))
    
    df <- data.frame(Time = pred, Probabilities = probs, Distribution = distribution)

    # Combine data
    combined_df <- rbind(combined_df, df)
}

# Generate a single plot with curves for all distributions
p <- ggplot(combined_df, aes(x = Time, y = Probabilities, color = Distribution)) +
    geom_line(linewidth=1.2) +
    labs(title = "Survival Curves for Various Distributions",
         x = "Survival Time",
         y = "Survival Probability") +
    theme_minimal()

# Print the plot
print(p)
```
Lognormal seems to be the better one, however, let's also compare the AIC and BIC.

```{r}
combined_scores <- data.frame(Distribution = character(), AIC = numeric(), BIC = numeric())


for(distribution in names(survreg.distributions)){
    reg_m <- reg_models[[distribution]]
    
    extracted_scores <- extract(
      reg_m,
      include.aic = TRUE,
      include.bic = TRUE
    )
    
    # Extract AIC and BIC names
    score_names <- extracted_scores@gof.names
    
    # Find indices of AIC and BIC in the names
    aic_index <- which(score_names == "AIC")
    bic_index <- which(score_names == "BIC")
    
    # Extract AIC and BIC scores
    aic <- extracted_scores@gof[aic_index]
    bic <- extracted_scores@gof[bic_index]
    
    combined_scores <- rbind(combined_scores, 
                             tibble(Distribution = distribution,
                                    AIC = aic,
                                    BIC = bic))
}
# Order the scores by AIC and BIC
combined_scores_ordered <- combined_scores[order(combined_scores$AIC, combined_scores$BIC), ]

# Print the combined scores data frame
print(combined_scores_ordered)
```

As we can see the models with `loggaussian` and `lognormal` distributions have lower AIC and BIC. Let's pick the `lognormal` one go on with it. We'll train a new model adding some of the variables to it. But first, we'll define the order to some of the factor variables.

```{r}
# Define the education order
ed_order <- c("Did not complete high school", "High school degree", "Some college", "College degree", "Post-undergraduate degree")

# Apply the education order to the respective variable
telco$ed <- factor(telco$ed, levels = ed_order)
```

Let's add gender all the columns to the model and then remove those that are not statistically significant to the model (we assume that the p-values of the models with only one covariate and the model with said covariate and some others are not very different).

```{r}
reg_f= survreg(surv_obj ~ region + age + marital + address + income + ed + retire + gender + voice + internet + forward + custcat,
               data=telco, dist="lognormal")
summary(reg_f)
```
As we can see, only the coefficients of `age`, `marital`, `voice`, `internet`, and `custcat` are statistically significant. Let's rebuild the model using only those. Let's also add `ed`, because, I suppose, it might also have some impact on the model (the p-value of some education levels are almost statistically significant).


```{r}
reg_f= survreg(surv_obj ~ age + marital +  voice + internet + custcat,
               data=telco, dist="lognormal")
summary(reg_f)
```
Now, as we have the model, let's have a look at AIC and BIC.

```{r}
extract(
    reg_f,
    include.aic = TRUE,
    include.bic = TRUE
)
```

As we can see the AIC and BIC are better than of the intercept-only model. Now, let's interpret the coefficients. Since we're using a `lognormal` distribution for the model, we'll need to exponentiate the coefficients returned by the model to understand the real effect of the covariates.

```{r}
exp(coef(reg_f))
```

## Report

As we can see the coefficients for `maritalUnmarried`, `voiceYes`, and `internetYes` are less than one. From this we can conclude that unot married individuals, as well as individuals using voice and/or internet services have less life time, that is, they are more prone to churn earlier compared to married, individuals not using voice and internet services, respectively.

On the contrary, `age` has a coefficient, which is greater than 1. This means as people get older, they are less prone to churn (the lifetime is longer). Same goes for the different customer categories; those using Plus service tend to have longer lifetime (2.4 times more), those using E-service and Total service are even less prone to churn (about 3 times longer lifetime), compared to individuals with Basic service.

## CLV
(Something is off here. I didn't have time to understand what the problem is before midnight. My next commits will solve the problem.)
```{r}
pred=predict(reg_f, type="response")

pred_data=data.frame(t(pred))[,0:24]

sequence = seq(1,length(colnames(pred_data)),1)
MM = 1300
r = 0.1

for (num in sequence) {
    pred_data[,num]=pred_data[,num]/(1+r/12)^(sequence[num]-1)
}
```


```{r}
pred_data$CLV=MM*rowSums(pred_data)
summary(pred_data$CLV)
```


```{r}
ggplot(pred_data,aes(x=CLV))+labs(title = "CLV Distribution")+
geom_histogram()
```


```{r}
telco$CLV = pred_data$CLV
ggplot(telco,aes(x=CLV, color=gender))+
labs(title = "CLV Density By Gender")+
geom_density()
```
